---
title: "cm007 Exercises: Practice with `dplyr`"
output: 
  html_document:
    keep_md: true
    theme: paper
---

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tsibble))
```


This worksheet contains exercises aimed for practice with `dplyr`. 


1. (a) What's the minimum life expectancy for each continent and each year? (b) Add the corresponding country to the tibble, too. (c) Arrange by min life expectancy.

```{r}
gapminder %>%  #Our dataset
  group_by(continent, year) %>% #Our dataset grouped by variables continent and year
  summarise(min_life = min(lifeExp), #Our grouped dataset summarized by minimum life expectancy; "summarise" is the same as "summarize".
            country = country[lifeExp == min_life]) %>% #then we store our country variable into a new arrangement. 
  arrange(min_life) #we finallly arrange our dataset by minimum life expectancy. 
```

2. Calculate the growth in population since the first year on record _for each country_ by rearranging the following lines, and filling in the `FILL_THIS_IN`. Here's another convenience function for you: `dplyr::first()`. 

```
mutate(rel_growth = FILL_THIS_IN) %>% 
arrange(FILL_THIS_IN) %>% 
gapminder %>% 
DT::datatable()
group_by(country) %>% 
```

```{r}
gapminder %>% #Our dataset
  group_by(country) %>% #Grouped by "country"
  arrange(year) %>% #Arranged by year
  mutate(rel_growth = pop - first(pop)) %>% #We add a new column called rel_growth where we calculate the growth of each year against the first year recorded.
  DT::datatable()
```


3. Determine the country that experienced the sharpest 5-year drop in life expectancy, in each continent, sorted by the drop, by rearranging the following lines of code. Ensure there are no `NA`'s. Instead of using `lag()`, use the convenience function provided by the `tsibble` package, `tsibble::difference()`:

```
drop_na() %>% 
ungroup() %>% 
arrange(year) %>% 
filter(inc_life_exp == min(inc_life_exp)) %>% 
gapminder %>% 
mutate(inc_life_exp = FILL_THIS_IN) %>% 
arrange(inc_life_exp) %>% 
group_by(country) %>% 
group_by(continent) %>% 
knitr::kable()
```

```{r}
gapminder %>% #Our dataset
  group_by(country) %>% #grouped by country
  arrange(year) %>%  #arranged by year
  mutate(inc_life_exp = difference(lifeExp)) %>% #add a new column called "inc_life_exp" which will be the diference of lifeExp
  drop_na() %>% #We leave out NA's (beacuse the first year cannot be compared)
  ungroup() %>% #Then we ungroup
  group_by(continent) %>% #And now we group again but this time by continent
  filter(inc_life_exp == min(inc_life_exp)) %>%  #then we filter the dataset to display only the country within each continent with the minimum increase of life expectancy recorded. 
  arrange(inc_life_exp) %>% #we arrange it by our new variable
  knitr::kable() 
  
```

Australia and Rwanda are both presented on the table, however notice that at the difference is quite big between the countries. 


